---
title: "02_TNBC_Spe_LC"
author: "Xinyue_Cui"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(cytomapper)
library(SpatialExperiment)
```

```{r}
TNBC <- read_csv("/Users/cuixinyue/Desktop/TNBC_analysis/Data/MIBI-TNBC_count_output.csv")
TNBC$sample_id <- as.factor(TNBC$sample_id )
TNBC$mm <- as.factor(TNBC$mm )
#TNBC$cell_id<- paste(TNBC$ImageNb, TNBC$cellLabelInImage, sep="_")
```
# Construct count matrix
##The aim of this section is to create a count matric to be used in creating the spatial experiment where the rows are the protein markers, and the columns are all cells in the output matrix. There are total of 34 markers in the MB_Xeno dataset, thus, the expected dimension of this count matrix is 34 rows by 179194 columns(cells).
```{r}
colnames(TNBC)
```

#TNBC
```{r}
# filter dataset of mouse c2m1 and c5m1
# MB_Xeno_1n2 <-MB_Xeno  %>% 
#   filter(mouse_id %in% c('c2m1','c5m1'))
# dim(MB_Xeno_1n2)
segments <- split(TNBC, TNBC$sample_id)
TNBC_1<- segments$`1`
dim(TNBC_1)
colnames(TNBC_1)
# extract the require columns(last 70 columns)[from 21 to ]
TNBC_1_require <-TNBC_1[, 21:64]
dim(TNBC_1_require )
colnames(TNBC_1_require )
# Transpose 
test_count_1<- t(TNBC_1_require )
dim(test_count_1)
rownames(test_count_1)
all.equal(rownames(test_count_1), colnames(TNBC_1_require))

```
#Application to create count matrix
```{r}
create_countMat <- function(dataset){
  # Extract the last 70 columns from the dataset(21 to 90 columns)
  dataset <- dataset[, 21:64]
  # De-select metal markers
  # Transpose the matrix to count matrix
  count_matrix <- t(dataset)
  return(count_matrix)
}
 test_countMat <- create_countMat(TNBC_1)
 all.equal(test_countMat, test_count_1)

 count_matrix <- create_countMat(TNBC)
 dim(count_matrix)
rownames(count_matrix)
```

```{r}

TNBC_countMat <- TNBC[, 21:64]%>% 
    select(-c("Au", "Background", "Fe", "Ca", "Ta", "Si", "P", "Na"))
count_matrix <- t(TNBC_countMat)
dim(TNBC_countMat)

colData <- TNBC %>% 
    select(-rownames(count_matrix))
dim(colData)
```
### for patient 4 (countMat)
```{r}
segments <- split(TNBC, TNBC$sample_id)
TNBC_4<- segments$`4`

TNBC_countMat4 <- TNBC_4[, 21:64]%>% 
    select(-c("Au", "Background", "Fe", "Ca", "Ta", "Si", "P", "Na"))
count_matrix4 <- as.data.frame(t(TNBC_countMat4))
class(count_matrix4)
dim(count_matrix4)
```

### for patient 12 (countMat)
```{r}
segments <- split(TNBC, TNBC$sample_id)
TNBC_12<- segments$`12`

TNBC_countMat12 <- TNBC_12[, 21:64]%>% 
    select(-c("Au", "Background", "Fe", "Ca", "Ta", "Si", "P", "Na"))
count_matrix12 <- as.data.frame(t(TNBC_countMat12))
class(count_matrix12)
dim(count_matrix12)
```

#Spatial information
```{r}
spatialCoords <- as.matrix(colData[, c("centroidX", "centroidY")])
head(spatialCoords)
dim(spatialCoords)
```

```{r}
# construct spatial experiment
spe <- SpatialExperiment(
  assay = count_matrix,
  colData = colData,
  spatialCoordsNames = c("centroidX", "centroidY")
  #spatialCoords = spatialCoords
)

assayNames(spe) <- "exprs"
spe
unique(spe$sample_id)
str(spe$sample_id)
```

```{r}
# adding prefixes for cell identity
colnames(spe) <- paste0("Cell_", 1:dim(assay(spe))[[2]])

# adding prefixes for sample identity
spe$sample_id[spe$sample_id %in% 1:9] <- (
  paste0("0", spe$sample_id[spe$sample_id %in% 1:9])
   )
spe$sample_id <- paste0("Sample_", spe$sample_id)
```

We want to verify if we have added the prefixes correctly.
```{r}
# checking sample identity
unique(spe$sample_id)
# checking cell identity
spe
```

Saving constructed spe object as RDS file.
```{r}
save(spe, file = ("/Users/cuixinyue/Desktop/TNBC_analysis/Output/RData/02_TNBC_spe_LC.rds"))
```

#Images and segmentation masks
```{r}
# imgData(spe)
# getImg(spe)
```



