---
title: "Diagnostic_of_LDA_TNBC"
author: "Xinyue_Cui"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Library
```{r warning=FALSE, message=FALSE}
#library(phyloseq)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)
#library(genefilter) 
library(rstan)
library(randomcoloR)
library(SpatialExperiment)
#library(DESeq2)
# library(here)
``` 
### Dataset
```{r}
load(file = here::here("Output","RData","02_TNBC_spe_LC.rds"))
spe_4
```
### Document-Term Matrix
```{r dtm}
dtm_4 <- as.data.frame.matrix(table(spe_4$sample_id, spe_4$mm)) %>% as.matrix()
dtm_4 <- as.matrix(table(spe_4$sample_id, spe_4$mm))
dim(dtm_4)
dim_names <- dimnames(dtm_4)
col_names <- colnames(dtm_4)
```

```{r}
K <- 5
alpha <- 0.8
gamma <- 0.8
x <- dtm_4
chain = 4
iter = 2000
```

### Posterior Sampling
```{r lda result, eval=FALSE}
source(here::here("R_notebooks","06_LDA_scripts", "06_get_lda_result.R"))

patients<- get_lda_result(K = 5, alpha = 0.8, gamma = 0.8,
                          dtm = dtm_4, iter = 4000, chains = 2,
                          file_fold = "05_LDA_multiChains", save_file = TRUE)
```
Load the existing LDA estimations.
```{r}
file_fold = "05_LDA_multiChains"
lda_file = "JABES_Endo_all_K_5_ite_2000_chains_2.RData"
file = here::here("Output", "RData", file_fold, lda_file)

load(file)
stan.fit
res_lda <- rstan::extract(
    stan.fit,
    permuted = TRUE,
    inc_warmup = FALSE,
    include = TRUE
  )
samples <-res_lda
dim(stan.fit)
```

Load the existing LDA estimations.
```{r}
source(here::here("R_notebooks", 
                  "06_LDA_scripts", "06_load_lda_result.R"))

file_fold = "05_LDA_multiChains"
lda_file = "JABES_Endo_all_K_5_ite_2000_chains_4.RData"

lda_result <- load_lda_result(
  file = here::here("Output", "RData", 
                   file_fold, lda_file)
)
```


```{r}
# theta estimation
theta <- samples$theta
dim(theta)

# beta estimation
beta <- samples$beta
dim(beta)
# lda_result
```

# Now, 'theta_indices' contains all indices of columns that include 'theta'

```{r}
# Extracting 'theta' samples, including warmup
# This will give you a matrix where rows are iterations (including warmup) and columns are chains
# theta_samples <- res[, , theta_index]


# ??rstan::extracts

# res[,,"theta[1,1]"]
# rownames(res[,,"chain:1"])
```

### Alignment matrix
```{r}
library(abind)
source(here::here("R_notebooks", "06_LDA_scripts", "06_alignmentMatrix.R"))

aligned_matrix <- alignmentMatrix(
  theta = theta,
  spe = spe_4,
  K = K,
  iter = iter,
  chain = chain,
  #cor_method = "cosine",
  SampleID_name = "sample_id"
)
aligned_matrix
```

### Align topic proportion

We align the topic proportion in each specimen across four chains.
```{r align-theta}
source(here::here("R_notebooks", "06_LDA_scripts", "06_thetaAligned.R"))

theta_aligned <- thetaAligned(
  theta = theta, 
  K = K, 
  aligned = aligned_matrix, 
  iter = iter,
  chain = chain,
  )
theta_aligned 
dim(theta_aligned)
```

### Align ASV proportion

We align the ASVs proportion in each topic across four chains.
```{r align-beta}
source(here::here("R_notebooks", "06_LDA_scripts", "06_betaAligned.R"))

beta_aligned <- betaAligned(
  beta = beta, 
  K = K, 
  aligned = aligned_matrix, 
  iter = iter,
  chain = chain,
  ) 
beta_aligned 
```
## Diagnostic after aligned
## Visualization

### Plot topic proportion

```{r plot-topic-prop, fig.width=20, fig.height=8}
source(here::here("R_notebooks", "06_LDA_scripts", "06_plotTopicProportion.R"))

p_topicProp <- plotTopicProportion(
  spe = spe_4,
  theta_aligned = theta_aligned,
  K = K,
  iter = iter,
  chain = chain,
  SampleID_name = "sample_id"
)
p_topicProp + theme(
  axis.title = element_text(size = 14),
  axis.text = element_text(size = 14),
  plot.title = element_text(size = 14))

#ggsave(filename = "Topic_Prop_5.png", path="/Users/henzhwang/Desktop/",
#       width = 15, height = 10)
```

```{r plot cellType Dist, fig.width=10, fig.height=10}
source(here::here("R_notebooks", 
                  "06_LDA_scripts", "06_plotCellTypeDistribution.R"))

p_cellTypeDist <- plotCellTypeDistribution(
  spe = spe_4,
  K = K,
  beta_aligned = beta_aligned,
  col_names_beta_hat = c("iterations", "Topic", "Cell.Type", "beta_h")
)

p_cellTypeDist
#ggsave(filename = "cellType_dist_5.png", path="/Users/henzhwang/Desktop/",
#       width = 15, height = 10)
```

# Diagnostic plot

## Diagnostic plot
Perform computation of effective sample size for each parameter
```{r}
source(here::here("R_notebooks", "06_LDA_scripts", "06_diagnosticsPlot.R"))

p_diag_5_2000 <- diagnosticsPlot(
  theta_aligned = theta_aligned,
  beta_aligned = beta_aligned,
  iter = iter,
  chain = chain,
) 
p_diag_5_2000

```

## Model Asseessement
### model assessment with max 
We want to perform goodness of fit test using the posterior estimates and observed data.
```{r}
# model assessment with max 
source(here::here("R_notebooks", "06_LDA_scripts", "06_modelAssessment_max.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_max <- modelAssessment_max(dtm = dtm_4,
                res = samples,
                warm_up_iter = NULL,
                iter = 2000,
                cellTypeIndexToPlot = c(1:16)
                )
p_model_assessment_max
```

### model assessment with mean
```{r}
# model assessment with mean 
source(here::here("R_notebooks", "06_LDA_scripts", "06_modelAssessment_mean.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_mean <- modelAssessment_mean(dtm = dtm_4,
                res = samples,
                warm_up_iter = NULL,
                iter = 2000,
                cellTypeIndexToPlot = c(1:16)
                )
p_model_assessment_mean
```
### model assessment with median
```{r}
# model assessment with median 
source(here::here("R_notebooks", "06_LDA_scripts", "06_modelAssessment_median.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_median <- modelAssessment_median(dtm = dtm,
                res = samples,
                warm_up_iter = NULL,
                iter = 2000,
                cellTypeIndexToPlot = c(1:16)
                )
p_model_assessment_median
```
## Trace plot Diagnostic after aligned
```{r}
source(here::here("R_notebooks", 
                  "06_LDA_scripts", "06_traceplot_aligned.R"))

tp <- traceplot_aligned(theta_aligned, 
                  num_chains = 4 , 
                  num_iterations_per_chain = 500)

```
## mean difference 

```{r}
x_sim <- samples$x_sim[,4, ] # iteration * tissues * topics


# Assuming samples$x_sim is your array and it's correctly indexed

# Extract the iterations for each chain
chain1 <- samples$x_sim[1:1000, 1, ]
chain2 <- samples$x_sim[1001:2000, 2, ]
chain3 <- samples$x_sim[2001:3000, 3, ]
chain4 <- samples$x_sim[3001:4000, 4, ]

# Now, bind these matrices together along the third dimension
all_chains <- array(dim = c(1000, 16, 4))
all_chains[, , 1] <- chain1
all_chains[, , 2] <- chain2
all_chains[, , 3] <- chain3
all_chains[, , 4] <- chain4

mean_across_chains <- apply(all_chains, c(1, 2), mean)
mean_across_chains_df <- as.data.frame(mean_across_chains)
colnames(mean_across_chains_df) <- colnames(x)
mean_across_chains_df


transposed_df <- as.data.frame(t(mean_across_chains_df))

# The row names of the transposed dataframe will be the column names of the original dataframe.
# To turn row names into a proper column, you might want to reset them:
colnames(transposed_df) <- rownames(transposed_df)
rownames(transposed_df) <- NULL
transposed_df$count = x_cellCount$count

transposed_df_minus_count <- transposed_df
count_column <- transposed_df_minus_count$count

transposed_df_minus_count <- transposed_df_minus_count[ , -which(names(transposed_df_minus_count) == "count")]
transposed_df_minus_count <- sweep(transposed_df_minus_count, 2, count_column, `-`)

colnames(transposed_df_minus_count) <- NULL

squared_df <- transposed_df_minus_count
squared_df[] <- lapply(squared_df, function(x) x^2)

# Calculate the row-wise sum of the squared values
row_sums_squared <- rowSums(squared_df)

```

## Diagnostic after aligned with warmup
```{r}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_traceplot_warmup.R"))

tp2 <- traceplot_warmup(theta, spe_4,
                              K=5,
                              warm_up_iter = NULL,
                              iterations = 2000,
                              chain = 4,
                              TissueID_name = "tissue_id")

```