---
title: "Tesselation_V2"
author: "Xinyue_Cui"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library

```{r library, warning=FALSE, message=FALSE}
library(rstan)
library(dplyr)
library(EBImage)
library(ggplot2)
library(spatstat)
library(randomcoloR)
library(SpatialExperiment)
library(sp)
library(tidyr)
library(tibble)
library(maptools)
library(spdep)
library(adespatial)
```

## Dataset

Load the SPE object.

```{r}
load(file = here::here("Output","RData","02_TNBC_spe_LC.rds"))
spe
```


## Posterior Sampling

Load posterior sampling results of 5 topics using 4 chains and 2000 iterations on the dataset.

```{r}
source(here::here("R_notebooks", 
                  "06_LDA_scripts", "06_LDA_analysis.R"))

TNBC_LDA_5_2000 <- LDA_analysis(
    spe = spe,
    SampleID_name = "sample_id",
    cellType_name = "mm",
    K = 5,
    alpha = 0.8,
    gamma = 0.8,
    warm_up_iter = NULL,
    iter = 2000,
    chain = 4,
    col_names_theta_all = c("iteration", "Sample", "Topic", "topic.dis"),
    col_names_beta_hat = c("iterations", "Topic", "Cell.Type", "beta_h"),
    stan_file_path = here::here("R_notebooks", "06_LDA_scripts", "06_lda.stan"),
    cellTypeIndexToPlot = c(1:16),
    load_file = TRUE,
    save_file = FALSE,
    file_default = TRUE,
    file_fold = "05_LDA_multiChains"
)
```
Load the existing LDA estimations.
```{r}
file_fold = "05_LDA_multiChains"
lda_file = "JABES_Endo_all_K_5_ite_2000_chains_4.RData"
file = here::here("Output", "RData", file_fold, lda_file)

load(file)
stan.fit
res <- rstan::extract(
    stan.fit,
    permuted = TRUE,
    inc_warmup = FALSE,
    include = TRUE
  )
samples <-res
dim(stan.fit)
```
## Cell Type Proportion in each Topics

```{r}
beta_hat <- TNBC_LDA_5_2000$beta_aligned
dim(beta_hat)
dimnames(beta_hat)
```

```{r}
library(sf)
library(terra)
library(spatstat)
library(tmap)
library(spdep)
```

```{r}
s_coor <- spatialCoords(spe) |>
  data.frame()

df <- bind_cols(
  sample_id = spe$sample_id, 
  cell_id = dimnames(assay(spe))[[2]], 
  cell_type = spe$mm,
  cell_size = spe$cellSize
  ) |>
  data.frame() |>
  cbind(s_coor) |>
  left_join(beta_median, by = "cell_type") |>
  mutate(tumor_not = ifelse(cell_type == "Other", 1, 0))

```


```{r}
sample_names <- df$sample_id %>%
  unique() %>%
  as.list()

ppp_list <- list()

for (i in seq_along(sample_names)) {
  spe_temp <- spe[ ,spe$sample_id == sample_names[[i]]]
  centroidX <- spatialCoords(spe_temp)[, 1]
  centroidY <- spatialCoords(spe_temp)[, 2]

  # Ensure df_temp is not empty
    ppp_temp <- ppp(
      centroidX,
      centroidY,
      window = owin(
        c(0, max(centroidX)),  # Correct range for x
        c(0, max(centroidY))   # Correct range for y
        ),
      marks = factor(spe_temp$mm)
      
    )

    ppp_list[[i]] <- ppp_temp
  }


names(ppp_list) <- sample_names
```

## Create a point pattern for each patient

```{r}
sam_names <- df$sample_id |>
  unique() |>
  as.list() 

ppp_lst <- list()

for(i in seq_along(sam_names)){
  df_temp <- dplyr::filter(df, sample_id == sam_names[i])

  ppp_temp <- ppp(
    x = df_temp$centroidX, 
    y = df_temp$centroidY,
    window = owin(
      c(0,max(df_temp$centroidX)),
      c(0,max(df_temp$centroidY))
      ),
    marks = select(
      df_temp, 
      cell_type,  
      cell_size, 
      tumor_not, 
      Topic_1, 
      Topic_2, 
      Topic_3, 
      Topic_4, 
      Topic_5
      )
    )

  ppp_temp$marks$tumor_not <- factor(
    ppp_temp$marks$tumor_not
    )
  ppp_lst[[i]] <- ppp_temp
  
}

names(ppp_lst) <- sam_names
```


```{r}
class(ppp_list[[1]])
A <- by(ppp_list[[1]], FUN = adaptive.density)
plot(A)

```
```{r}
ppp_4<-ppp_list[[4]]
dir_tess_4 <- dirichlet(ppp_4)
plot(dir_tess_4)
#plot(dirichlet(ppp_list[[4]]))
plot(ppp_4, which.marks = c("cell_size", "tumor_not"), cols = c("red", "blue")) 

```
```{r}
solist(ppp_list[[4]], ppp_list[[11]])

L <- as.anylist(lapply(ppp_list[c(4,11)], Lest) )
plot(L, equal.scales=TRUE)
```

```{r}
polygon_data <- extract_polygons(dir_tess_4)
dir_poly_4 <- convert_to_SpatialPolygons(polygon_data)
plot(dir_poly_4)

```
### Tessellation

```{r}
dir_tess_p4 <- dirichlet(ppp_4)
dirichlet(ppp_4,10)
plot(dir_tess_p4)
```
```{r}
ppp_4$tile
# Plot the first 6 tiles
plot(dir_tess_p4 , which = 1:6)
library(maptools) 
library(spatstat.geom)
library(spatstat.core)
library(sf)
library(spatstat)

tiles_list <- tess2SP(dir_tess_p4)
```
```{r}
voronoi_tess <- dirichlet(ppp_4)
```

### Convert tiles to polygons

```{r}
dir_poly_p4 <- as(dir_tess_p4, "SpatialPolygons")
plot(dir_poly_p4)
```




```{r}
library(spdep)
library(sf)
library(sp)
library(lwgeom)

## queen's distance
dir_poly_nb_20 <- poly2nb(dir_poly_20)
dir_poly_net_20 <- nb2lines(dir_poly_nb_20, coords = coordinates(dir_poly_20))

dir_poly_lw1_20 <- nb2listw(dir_poly_nb_20, zero.policy = TRUE, style = "W")
print(dir_poly_lw1_20)

## rook's distance
dir_poly_nb_rook_20 <- poly2nb(dir_poly_20, queen = FALSE)
dir_poly_net_rook_20 <- nb2lines(dir_poly_nb_rook_20, coords = coordinates(dir_poly_20))

dir_poly_lw1_rook_20 <- nb2listw(dir_poly_nb_rook_20, zero.policy = TRUE, style = "W")
# print.listw(dir_poly_lw1_rook_s1, zero.policy = TRUE)
```

### global Moran's I

```{r}
df$centroid_x <- spatialCoords(spe)[, 1]
df$centroid_y <- spatialCoords(spe)[, 2]

# moran.test(spa_compa_s1[spa_compa_s1$tumor_not == 0,]$centroid_x, dir_poly_lw1_s1, zero.policy = TRUE)
# moran.test(spa_compa_s1[spa_compa_s1$tumor_not == 0,]$centroid_y, dir_poly_lw1_s1, zero.policy = TRUE)
```

```{r}
cut_ppp_20 <- cut(ppp_20, dir_tess_20, labels = TRUE)
plot(cut_ppp_20)
plot(quadratcount(cut_ppp_s1))
split_ppp_20 <- split(cut_ppp_20, dir_tess_20)

plot(split_ppp_20[1:30], use.marks = FALSE)
plot(split_ppp_20[31:60], use.marks = FALSE)
plot(split_ppp_20[61:90], use.marks = FALSE)
plot(split_ppp_20[91:120], use.marks = FALSE)
```

### Convert polygons to neighborhood

The function poly2nb allows to define neighborhood when the sampling sites are polygons and not points (two regions are neighbors if they share a common boundary).

```{r}
xx <- poly2nb(dir_poly_20)
plot(dir_poly_20, border = "grey")
plot(xx, coordinates(dir_poly_20),  pch = 20, col = "red")
title(main="Neighborhood for polygons")
```

```{r}
xx[[1]]
```
```{r}
listwgab <- nb2listw(xx)
listwgab

names(listwgab)

listwgab$neighbours[[1]]

listwgab$weights[[1]]
```
```{r}
mem.gab <- mem(listwgab)
mem.gab
```

```{r}
barplot(attr(mem.gab, "values"), 
        main = "Eigenvalues of the spatial weighting matrix", cex.main = 0.7)
```
```{r warning=FALSE, message=FALSE}

dir_poly_20$MEM1 <- mem.gab[, "MEM1"]
dir_poly_20$MEM2 <- mem.gab[, "MEM2"]
dir_poly_20$MEM3 <- mem.gab[, "MEM3"]

tmap::tm_shape(dir_poly_20) +
  tmap::tm_polygons(col = "MEM1")
  
tmap::tm_shape(dir_poly_20) +
  tmap::tm_polygons(col = "MEM2")

tmap::tm_shape(dir_poly_20) +
  tmap::tm_polygons(col = "MEM3")

```


### Test each MEM

Default `moran.randtest` - tests against the alternative hypothesis of positive autocorrelation.

```{r eval = FALSE}
moranI <- moran.randtest(mem.gab, listwgab, 99)
moranI

# save(moranI, file = "moranI.RData")
# load("moranI.RData")
```

```{r}
head(attr(mem.gab, "values") / moranI$obs)

signi <- which(moranI$pvalue < 0.05)
length(signi)
```

```{r}
signi[100]
dir_poly_20$MEM100 <- mem.gab[, signi[100]]


tmap::tm_shape(dir_poly_20) +
  tmap::tm_polygons(col = "MEM100")
  
```

