---
title: "TNBC Linear Mixed Model"
author: "Xinyue_Cui"
date: "2023-10-07"
output: html_document
---
# packages
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(cytomapper)
library(ggplot2)
library(formattable)
library(magrittr)
library(lme4)
library(car)
```

# The dataset TNBC

```{r}
TNBC <- read_csv("/Users/cuixinyue/Desktop/TNBC_analysis/Data/MIBI-TNBC_count_output.csv")
TNBC$sample_id <- as.factor(TNBC$sample_id )
TNBC$STAGE <- as.factor(TNBC$STAGE)
TNBC$mm <- as.factor(TNBC$mm)
TNBC$cell<- paste(TNBC$ImageNb, TNBC$cellLabelInImage, sep="_")
```

```{r}
table(TNBC$sample_id)
dim(TNBC)
table(TNBC$STAGE)
str(TNBC)
```

create a contingency table that displays the frequency counts of the combinations between the sample_id and STAGE variables from the TNBC dataset
```{r}
length(table(TNBC$sample_id,TNBC$STAGE))
```
```{r}
# get a distribution plot of CD45
ggplot(TNBC, aes(x = CD45)) +
  geom_density(fill = "blue", alpha = 0.5)  +
  labs(title = "Density Plot of logistic transformation of dsDNA_asinh_znorm ",
       x = "Variable Value",
       y = "Density")
```
Interpret of LMM
Intercept :the estimated mean value for the group of  Ct treatment. 
Treatement_idAraC_D2: the difference in the response between AraC_D2 and Ct.
Treatement_idAraC_D9: the difference in the response between AraC_D9 and Ct.

```{r}
head(TNBC)
TNBC.lme2 <- lmer(CD45~ Survival_days_capped + (Survival_days_capped| sample_id), data= TNBC)
summary(TNBC.lme2 )
ranef(TNBC.lme2)
fixef(TNBC.lme2)
```
# Build a data set for graph
```{r}
subset_TNBC <- TNBC %>%
  select(Survival_days_capped, sample_id,CD45)
```

```{r}
ggplot(subset_TNBC ,aes(x=Survival_days_capped,y=CD45))+geom_point()+facet_wrap(~sample_id, ncol = 8, strip.position = "top")+
  theme_minimal() 
```

```{r}
df <- data.frame(ranef(TNBC.lme1))
#STAGE10 is the intercept
STAGE10 <- data.frame(
  sample_id = df$grp[df$term == "(Intercept)"],
  type = "mixed model",
  STAGE1 = "STAGE10",
  CD45 = df$condval[df$term == "(Intercept)"]
)
slope <- data.frame(
  sample_id = rep(unique(subset_TNBC$sample_id), each = 8),
  STAGE1 = rep(unique(subset_TNBC$STAGE1),times= 39),
  type = "mixed model"
)
```

```{r}
merged <- slope %>%
  left_join(df, by = c("sample_id" = "grp", "STAGE1" = "term"))
merged_data <- merged  %>%
  left_join(STAGE10, by = c("sample_id" = "sample_id"))

slope_final <- data.frame(
sample_id =  merged_data$sample_id,
type= "mixed model",
STAGE1 = merged_data$STAGE1.x,
CD45 = merged_data$condval+merged_data$CD45
)

val_CD45 <- na.omit(rbind(STAGE10,slope_final), cols = "CD45")
```

```{r}
#fixed 

fixef(TNBC.lme1)
df_fix <- data.frame(fixef(TNBC.lme1))

fixed <- data.frame(
  sample_id = rep(unique(subset_TNBC$sample_id), each = 8),
  type = "all",
  STAGE1 = rep(unique(subset_TNBC$STAGE1),times= 39),
  CD45 = rep(df_fix$fixef.TNBC.lme1.,each = 39)

)

# Combine the data
obs <- subset_TNBC %>%
  select(sample_id,type, STAGE1 ,CD45)

plot_data <-rbind(obs,val_CD45,fixed)
plot_data$STAGE <- str_remove(plot_data$STAGE1, "STAGE")
plot_data
```

```{r}
#plot
ggplot(plot_data ,aes(x=STAGE,y=CD45,color = type))+geom_point()+facet_wrap(~sample_id, ncol = 8, strip.position = "top")+
  theme_minimal() 
```

