---
title: "12_Spatial_LDA_patient_4"
author: "Xinyue_Cui"
date: "2024-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , packages}
library(rstan)
library(dplyr)
library(EBImage)
library(ggplot2)
library(spatstat)
library(randomcoloR)
library(SpatialExperiment)
library(sp)
library(tidyr)
library(tibble)
library(maptools)
library(spdep)
library(adespatial)
```

# Apply Spatial LDA for MULTI
## Load the existing LDA estimations.
```{r}
file = "/Users/cuixinyue/Desktop/TNBC_analysis/Output/RData/Spatial_LDA_results/slda_lda_tile_4000_chain2_patient4.RData"

load(file)
fit_12 
res_12<- rstan::extract(
    fit_12 ,
    permuted = TRUE,
    inc_warmup = FALSE,
    include = TRUE
  )

apply(res_12$theta[,3,], 2, median)
dim(res_12$theta[,3,])
```
```{r}
dim(res_12$theta[3001:4000,,])
median(res_12$theta[,1,1]) # 0.0002692714
median(res_12$theta[,1,2]) # 0.2375964
median(res_12$theta[,1,3]) # 0.006216302
median(res_12$theta[,1,4]) # 0.01350381
median(res_12$theta[,1,5]) # 0.4735648
dim(est_doc_topic_prop_lda)

est_doc_topic_prop_lda[1,]/sum(est_doc_topic_prop_lda[1,])

res_theta <- res_12$theta[3001:4000,,]
```


```{r}

est_doc_topic_prop_lda <- apply(
  res_theta, 
  MARGIN = c(2, 3), 
  median
  )
est_doc_top_lda <- paste0(
  "topic",
  max.col(est_doc_topic_prop_lda)
  )

df_doc_top_lda <- as.data.frame(est_doc_topic_prop_lda)
df_doc_top_lda $tile <- seq_len(nrow(df_doc_top_lda ))
df_doc_top_lda <- df_doc_top_lda %>%
  rename(
    topic1_slda = V1,
    topic2_slda = V3,
    topic3_slda = V5,
    topic4_slda = V2,
    topic5_slda = V4
  )

df_doc_top_lda$tile <- as.factor(df_doc_top_lda$tile)

# Now try the join again
S12_info_slda <- S12_info %>% left_join(df_doc_top_lda, by = c("tile_align" = "tile"))
S12_info_slda

```
```{r normalized_theta_proportion}
est_doc_topic_prop_lda <- apply(
  res_theta, 
  MARGIN = c(2, 3), 
  median
  )
tiles = 1000
est_doc_topic_prop_slda <- array(dim=c(1000, 5))
for (i in 1:tiles) {
  est_doc_topic_prop_slda[i,] <-est_doc_topic_prop_lda[i,]/sum(est_doc_topic_prop_lda[i,])
}

est_doc_top_slda <- paste0(
  "topic",
  max.col(est_doc_topic_prop_slda)
  )

df_doc_top_slda <- as.data.frame(est_doc_topic_prop_slda)
df_doc_top_slda $tile <- seq_len(nrow(df_doc_top_slda ))
df_doc_top_slda <- df_doc_top_slda %>%
  rename(
    topic1_slda = V1,
    topic2_slda = V3,
    topic3_slda = V5,
    topic4_slda = V2,
    topic5_slda = V4
  )

df_doc_top_slda$tile <- as.factor(df_doc_top_slda$tile)

# Now try the join again
S12_info_slda_v2 <- S12_info %>% left_join(df_doc_top_slda, by = c("tile_align" = "tile"))
S12_info_slda_v2 
```

## Spatial LDA topic plots
#### Plot function for Spatial LDA

```{r}
topic_plot <- function(info_data, topicprop) {
  p <- ggplot(info_data, aes_string(x = "centroidX", y = "centroidY", color = topicprop)) +  # use aes_string for variable names as strings
    geom_point(size = 1, alpha = 1) +  # Adjust point size and transparency
    scale_color_gradient(low = "white", high = "red", 
                         limits = c(0, 1),  # Ensure the limits are set from 0 to 1
                         oob = scales::squish) +  # Squish values outside the range into the limits
    theme_minimal() +
    labs(title = 'Spatial Plot of Cells', x = 'Centroid X', y = 'Centroid Y') +
    theme(legend.position = "right")  # Adjust legend position if needed

  return(p)
}
```

```{r}
topic_plot(S12_info_slda_v2,"topic1_slda")
topic_plot(S12_info_slda_v2,"topic2_slda")
topic_plot(S12_info_slda_v2,"topic3_slda")
topic_plot(S12_info_slda_v2,"topic4_slda")
topic_plot(S12_info_slda_v2,"topic5_slda")
```
```{r}
topic_plot(S12_info_slda,"topic1_slda")
topic_plot(S12_info_slda,"topic2_slda")
topic_plot(S12_info_slda,"topic3_slda")
topic_plot(S12_info_slda,"topic4_slda")
topic_plot(S12_info_slda,"topic5_slda")
```


## Topic Align
```{r}
slda_theta<- res_12$theta
slda_phi<- res_12$phi
dim(res_12$phi)
dim(res_12$theta)
```
#### Alignment Matrix

```{r}
  theta = slda_theta
  tile_long_doc = tiles_long_12
  #cor_method = "cosine",
  TileID_name = "tile_id"
  K = 5
  iter = 4000
  chain = 2
  warm_up_iter <- 3000

 Chain <- Topic <- topic.dis <- NULL
  # theta is a 3 dimensional array (iterations * samples * topic)
  
  # determine the iteration used in posterior sampling (subtract warm up iterations)
  if (is.null(warm_up_iter)) {
    iterUse = iter / 2
  } else {
    iterUse = iter - warm_up_iter
  }
  
dimnames(theta)[[2]] <- unique(tile_long_doc$tile_id)
  ## acquire topic number
dimnames(theta)[[3]] <- c(paste0("Topic_", seq(1,K)))
  theta_all = reshape2::melt(theta)
  colnames(theta_all) = c("iteration", "Tile", "Topic", "topic.dis")
  theta_all$Chain = paste0("Chain ", rep(seq(1, chain), each = (iterUse)))
  
  theta_all$Topic = factor(theta_all$Topic)
  theta_all$Chain = factor(theta_all$Chain)
  theta_all$Tile = as.character(theta_all$Tile)

aligned <- matrix(nrow = K, ncol = chain)
aligned[, 1] <- seq(1, K)
corrTop <- numeric()

  for(j in 1:K) { #Topic of first chain
    chains <- lapply(as.list(1:chain), function(x){
      for(top in 1:K){
        corrTop[top] <- cor(
          theta_all |> dplyr::filter(Chain == "Chain 1") |> dplyr::filter(Topic == paste0("Topic_", j)) |> dplyr::select(topic.dis),
          theta_all |> dplyr::filter(Chain == paste0("Chain ",x)) |> dplyr::filter(Topic == paste0("Topic_", top)) |> dplyr::select(topic.dis))
      }
      return(which(corrTop == max(corrTop)))
    })
    
    aligned[j, 2:chain] <- unlist(chains)[2:chain]
  }

aligned_matrix <- aligned
```

```{r}
# aligned_matrix[5, 4] <- 3
aligned_matrix
```

#### Align topic proportion

We align the topic proportion in each specimen across four chains.
```{r align-theta}
  theta = slda_theta
  tile_long_doc = tiles_long_4
  
 if (is.null(warm_up_iter)) {
    iterUse = iter / 2
  } else {
    iterUse = iter - warm_up_iter
  }
theta <- aperm(theta, c(1,3,2))

  theta_chain <- list()
  theta_chain[[1]] <- theta[1:(iterUse), ,]
  theta_chain[[1]] <- theta_chain[[1]][, aligned[,1],]
  
  theta_aligned <- theta_chain[[1]]

  for(ch in 2:chain){
    theta_chain[[ch]] <- theta[((ch-1)*(iterUse)+1):(ch*(iterUse)),,]
    theta_chain[[ch]] <- theta_chain[[ch]][,aligned[,ch],]
    
    theta_aligned <- abind::abind(theta_aligned, theta_chain[[ch]], along = 1)
  }

  # theta_aligned <- abind(theta_chain[[1]], theta_chain[[2]], along = 1)
  # for(ch in 3:chain){
  #   theta_aligned <- abind(theta_aligned, theta_chain[[ch]], along = 1)
  # }
  # switch back samples and Topic dimension in array
  theta_aligned <- aperm(theta_aligned, c(1,3,2))


slda_theta_aligned <- theta_aligned 
dim(slda_theta_aligned)
```
#### Align ASV proportion
We align the ASVs proportion in each topic across four chains.
```{r align-beta}
# source(here::here("R_notebooks", "06_LDA_scripts", "06_betaAligned.R"))
beta <- slda_phi

 beta_chain <- list()
  beta_chain[[1]] <- beta[1:(iterUse), ,]
  beta_chain[[1]] <- beta_chain[[1]][, aligned[,1], ]
  
  beta_aligned <- beta_chain[[1]]

  for(ch in 2:chain){
    beta_chain[[ch]] <- beta[((ch-1)*(iterUse)+1):(ch*(iterUse)),,]
    beta_chain[[ch]] <- beta_chain[[ch]][,aligned[,ch],]
    
    beta_aligned <- abind::abind(
      beta_aligned,
      beta_chain[[ch]],
      along = 1)
  }
  
slda_beta_aligned <- beta_aligned
dim(slda_beta_aligned)
```
```{r}
source(here::here("R_notebooks", 
                  "06_LDA_scripts", "06_plotCellTypeDistribution.R"))
K = 5
p_cellTypeDist <- plotCellTypeDistribution(
  spe = spe_4,
  K = K,
  beta_aligned = slda_beta_aligned,
  col_names_beta_hat = c("iterations", "Topic", "Cell.Type", "beta_h")
)

p_cellTypeDist
```

## diagnostics

#### R_hat and ESS

```{r}
source(here::here("R_notebooks", "06_LDA_scripts", "06_diagnosticsPlot.R"))

p_list <-diagnosticsPlot(
  theta_aligned = slda_theta_aligned ,
  beta_aligned = slda_beta_aligned ,
  warm_up_iter = 3000,
  iter = 4000,
  chain = 2
) 

p_list
```

```{r}
slda_theta <- slda_theta_aligned
slda_phi<- slda_beta_aligned
# for theta
 Rhat_theta <- matrix(
    nrow = dim(slda_theta )[2],
    ncol = dim(slda_theta)[3]
  )
  
  ESS_bulk_theta <- matrix(
    nrow = dim(slda_theta)[2],
    ncol = dim(slda_theta)[3]
  )
  
  for(sam in 1:dim(slda_theta)[2]){
    for(top in 1:dim(slda_theta)[3]){
      sims_theta <- matrix(
        slda_theta[ ,sam , top],
        nrow = (iterUse),
        ncol = chain,
        byrow = FALSE
      )
      Rhat_theta[sam, top] <- rstan::Rhat(sims_theta)
      ESS_bulk_theta[sam, top] <- rstan::ess_bulk(sims_theta)
    }
    
  }

Rhat_theta <- as.vector(Rhat_theta)
ESS_bulk_theta <- as.vector(ESS_bulk_theta)


# For phi
  
  Rhat_phi <- matrix(
    nrow = dim(slda_phi)[2],
    ncol = dim(slda_phi)[3]
  )
  ESS_bulk_phi <- matrix(
    nrow = dim(slda_phi)[2],
    ncol = dim(slda_phi)[3]
  )
  
  
  for(top in 1:dim(slda_phi)[2]){
    for(fea in 1:dim(slda_phi)[3]){
      sims_phi <- matrix(
        slda_phi[ , top, fea],
        nrow = (iterUse),
        ncol = chain,
        byrow = FALSE)
      Rhat_phi[top, fea] <- rstan::Rhat(sims_phi)
      ESS_bulk_phi[top, fea] <- rstan::ess_bulk(sims_phi)
      
    }
    
  }
  
  Rhat_phi <- as.vector(Rhat_phi)
  ESS_bulk_phi <- as.vector(ESS_bulk_phi)
  

# combine the R_hat and ESS_bulk
Rhat <- c(Rhat_theta, Rhat_phi)
ESS_bulk <- c(ESS_bulk_theta, ESS_bulk_phi)

#plot for R_hat
# R hat ~ 1.05
  p_rhat <- ggplot2::ggplot(
    data.frame(Rhat = Rhat)
  ) +
    ggplot2::geom_histogram(
      aes(x = Rhat),
      fill = "lavender",
      colour = "black",
      bins = 100
    ) +
    ggplot2::theme(
      plot.title = element_text(hjust = 0.5)
    )  +
    ggplot2::theme_minimal(base_size = 20) +
    ggplot2::xlab("")
  
  
  
  
  # ESS bulk and ESS tail at least 100 per Markov Chain in order to be reliable and indicate that estimates of respective posterior quantiles are reliable
  
  p_ess_bulk <- ggplot2::ggplot(
    data.frame(ESS_bulk = ESS_bulk)
  ) +
    ggplot2::geom_histogram(
      aes(x = ESS_bulk),
      fill = "lavender",
      colour = "black",
      bins = 100
    ) +
    ggplot2::theme(
      plot.title = element_text(hjust = 0.5)
    )   +
    ggplot2::theme_minimal(
      base_size = 20
    ) +
    ggplot2::xlab("")

list(p_ess_bulk, p_rhat)
```
#### Traceplot
```{r}

```

## model assesment
#### dtm
```{r}
dtm <- as.data.frame.matrix(table(spe$sample_id, spe$mm)) %>% as.matrix()
dtm <- as.matrix(table(spe$sample_id, spe$mm))
dim(dtm)
dim_names <- dimnames(dtm)
col_names <- colnames(dtm)

```
#### sim
```{r}
dim(slda_beta_aligned)
result_sim <- array(dim=c(2000, 1000, 16))

slda_phi <-slda_beta_aligned
slda_theta <- slda_theta_aligned
tiles=1000
# Perform the multiplication using a for loop for the dimensions
for (i in 1:iterUse) {
  for (j in 1:tiles) {
    for (k in 1:16) {
      # Multiply the corresponding elements and sum over the common dimension
      result_sim [i, j, k] <- sum(slda_theta[i, j, ] * slda_phi[i, , k])
    }
  }
}

# Check the structure of the created 'sim' array
dim(result_sim)

result_sim <- result_sim*length(ppp_12$cluster)
```
#### Assemesent plot
```{r}
patient_id_index = 12
cellTypeIndexToPlot = c(1:16)
value <- Var2 <- NULL
  
  # determine the iteration used in posterior sampling (subtract warm up iterations)
  if (is.null(warm_up_iter)) {
    iterUse = iter / 2
  } else {
    iterUse = iter - warm_up_iter
  }
  
  x <- dtm

    x_cellCount <- data.frame(
    Var1 = rep(
      "x_mean_obs",
      dim(x)[2]
    ),
    Var2 = colnames(x),
    count = x[patient_id_index,] ## change here for different patient
  )
   x_cellCount <- x_cellCount[cellTypeIndexToPlot, ]
  # draws from posterior predictive distribution
  ### x_sim <- result$theta[1:iterUse, , ] # iteration * tissues * topics
  mean_all <- apply(result_sim[1, , ], 2, mean)
  for (i in 2:iterUse){
    mean_x_sim_i <- apply(
      result_sim[i, ,], 
      2,
      mean)
    mean_all <- rbind(mean_all, mean_x_sim_i)
  }
  
  rownames(mean_all) <- c(
    paste0(
      "x_mean_rep",
      seq(1, iterUse)
    )
  )
  
  colnames(mean_all) <- colnames(x)
  
  mean_all <- mean_all[, cellTypeIndexToPlot]
  mean_all_long <- reshape2::melt(mean_all)
  
  p_hist <- ggplot2::ggplot(
    data = mean_all_long
  ) +
    ggplot2::geom_histogram(
      aes(
        x = value,
        group = Var2
      ),
      color = "#0072B2",
      fill = "#0072B2",
      bins = 50) +
    ggplot2::xlab("mean")+
    
    ggplot2::facet_wrap(~Var2, nrow = 4, scales = "free_x") +
    
    ggplot2::geom_vline(
      data = x_cellCount,
      aes(xintercept = count),
      color = "#CC79A7"
    ) +
    ggplot2::theme_update(
      text = element_text(size = 8)
    )
  
p_hist 
```
#### Compare the RMSE
```{r}
mean_all_long_diff <- mean_all_long %>% left_join( x_cellCount, by = "Var2")
mean_all_long_diff$diff <- (mean_all_long_diff$value - mean_all_long_diff$count)^2

library(dplyr)

# Assuming your data frame is named mean_all_long_diff
sums_by_type <- mean_all_long_diff %>%
  group_by(Var2) %>%
  summarise(sum_diff = sum(diff))
sums_by_type$sum_sqrt_slda <- sqrt(sums_by_type$sum_diff)
sums_by_type


# df_RMSE <- subset(sums_by_type, Var2 != "Other")
```
```{r}
df_RMSE <- sums_by_type %>% left_join(sums_by_type_lda , by = "Var2")
df_RMSE 
p <- ggplot(df_RMSE, aes(x = sum_sqrt_lda.x, y = sum_sqrt_lda.y)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "blue") +
  labs(x = "slda_RMSE", y = "lda_RMSE", title = "Scatter plot of model Assessment of Slda vs lda") +
  theme_minimal()

p
```



